Manuscript

### 2022-09-04 Manuscript.tex
update manuscript

\title{Manuscript}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Jeff S. Wesner\textsuperscript{4}}
\date{30 June, 2022}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\setstretch{1}
\textsuperscript{1} Colorado Mesa University\\
\textsuperscript{2} Great Lakes Research Center, Michigan Technological
University, Houghton, MI USA\\
\textsuperscript{3} Louisiana Universities Marine Consortium, Chauvin,
LA USA\\
\textsuperscript{4} Dept. of Biology, University of South Dakota,
Vermillion, SD, USA

@@

Body size distributions are a fundamental characteristic of communities.
The remarkable consistency of these relationships across spatiotemporal
scales and habitat types has led them to be recommended as a
``universal'' indicator of ecological status (Petchey and Belgrano
2010)(Petchey and Belgrano 2010). Size metrics have commonly been used
in marine systems, and are increasingly being applied to assess the
condition of freshwater ecosystems (Martinez et al.~2016, Pomeranz
2018)(\textbf{martinez2016?}, \textbf{pomeranz2018?}).

@@

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
on empirical body sizes of stream benthic communities reported in
Pomeranz et al.~(2020)(\textbf{pomeranz2020?}). Our results are not
dependent on the range of body sizes, and we show the results of other
ranges in the supplemental information.

For the main analysis, presented here, we sampled \(n = 1000\) body
sizes from distributions described by five different \(\lambda\)'s:
(-1.5, -1.75, -2.00, -2.25, -2.5). The supplemental information contains
results when the value of n is varied (or maybe include in main?). Each
value of \(\lambda\) was assumed to come from a community across a
hypothetical environmental gradient \(X\). Values of \(X\) for each
community are uniformly distributed from -1 to 1. We repeated the data
simulation 1000 times (reps). The main results presented here were not
dependent on the range of \emph{x}-values or the number of sites
(supplemental information).

\hypertarget{sample-size-n}{%
\subsubsection{Sample size, n}\label{sample-size-n}}

The number of observations in our simulations may bias the results.
Therefore, we repeated the simulations described above, but varied the
sample size \(n\). We tested values of
\(n = 200, 500, 1000, 5000, 10 000\). (originally had n=100, but thre
errors based on random seed. Bumped it to 200 to try and get consistent
error-free runs)

@@

It is possible that our sampling or simulation framework could impact
the results obtained, possibly by inflating the type I error rate. In
order to assess whether or not our simulations are robust to this, we
performed independent random samples from a bounded power law for five
communities across a hypothetical gradient as described above, but we
set \(\lambda = -2\) for all five communities. Hence, we are simulating
a scenario where the size spectra relationship is invariant to our
hypothetical environmental gradient, and we should expect the estimated
relationship coefficient \(\beta_1 = 0\).

@@

For the main analysis, we were interested in determining whether or not
the methods were able to detect relatively coarse changes to size
spectra parameters, specifically, a change of 0.5 units across each
value of \(X\), or an absolute change or 1.0 units. However, many
seminal works on variation in size spectra relationships in freshwater
are on a much finer scale (on the order 0f 0.1 to 0.25 units).
Therefore, we performed simulations as described above, but varied
\(\lambda\) from -1.9 to -2.1 across the hypothetical environmental
gradient. This equates to a know relationship coefficient of
\(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
change of 0.2 units.

@@

For each sample of \(M\) and each of the 1000 replicates, we estimated
the exponent \(\lambda\) using MLE methods modified from the the
\texttt{sizeSpectra} package (Edwards et al.~2020). In addition, we use
two common binning methods to estimate the OLS slope parameter in
log-log space. For the first binning method, we created 6 equal
logarithmic bins covering the range of body sizes. The count in each bin
was normalized by dividing by the bin width. This method has been used
by (refs here, I think a lot of Woodward's group uses this\ldots)
Throughout the manuscript, the normalized equal logarithmic binning
method will be referred to as ELBn. The second method was similar to
ELBn, but bins of Log\textsubscript{2} widths are used. The count in
each bin is normalized in the same way. This method has been used by
(Pomeranz et al.~2018, I \textbf{think} McGarvey \textasciitilde2018,
Blanchard (log 2, ?normalized?))(McGarvey et al. 2019,
\textbf{pomeranz2018?}) and is referred to as the Normalized Abundance
Spectrum (NAS).

@@

For each simulation replicate, we estimated how the size spectra
parameters varied across the hypothetical gradient. Simple OLS
regression were conducted in the form
\(\lambda_{estimate} ~ \beta_0 + \beta_1 * X\). The distribution of the
relationship coefficient, \(\beta_1\), were plotted compared to the
known relationship. Likewise, the confidence interval of \(\beta_1\) was
assessed to determine the proportion of estimates which contained the
true value of the known relationship.

@@

We re-analyze two datasets of benthic macroinvertebrate communities from
stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage stress gradient. Details of the sample
collection and processing can be found in Pomeranz et
al.~(2018)(\textbf{pomeranz2018?}). Briefly, all individuals from each
sample were identified to the lowest practical taxonomic unit and body
lengths were measured using image processing software from photos taken
with a camera mounted to a dissecting microscope. Body mass was
estimated using published length weight regressions.

The second data set was from the wadeable stream sites of the National
Ecological Observatory Network (NEON; \textbf{data product
XXX})(National Ecological Observatory Network (NEON) 2022). These sites
are located across a wide temperature gradient in the United States,
from Puerto Rico to Alaska. Quantitative macroinvertebrate samples were
collected using the most appropriate method based on the local habitat.
All individuals were identified and had their body lengths measured, and
body mass was estimated using published length weight regressions. This
data has been analyzed previously using size spectra methods as
described in Pomeranz et al.~(2022)(Pomeranz et al. 2022). Detailed
methods of the sampling collection and data processing methods can be
found on the NEON website
({[}\url{https://www.neonscience.org/data-collection/macroinvertebrates}{]},
or macroinvertebrate DPI pubs).

@@

All methods performed reasonably well in detecting the known
relationship across the hypothetical environmental gradient. The
regressions using the MLE method found a significant relationship 100\%
of the time, while the ELBn and NAS method found significant
relatoinships 95.8\% and 96.9\% of the time, respectively.

The confidence intervals for the relationship coefficients (\(\beta_1\))
for both the MLE and ELBn method contained the true value of the known
relationship across the gradient \(\ge\) 95\% of the time, whereas the
confidence interval for the NAS method only had the true value 83.1\% of
the time. Despite having similar accuracy, the width of the CIs for the
ELBn method were more than 3 times that of the MLE method (ELBn\_CI\_ =
0.386 \(\pm\) 0.20; MLE\_CI\_ = 0.121 \(\pm\) 0.056). The CI's for the
NAS were slightly smaller than the ELBn method, but still \(>\) 2.5
times as wide as the MLE method (NAS\_CI = 0.339 \(\pm\) 0.172).

On average, the relationship across the gradient was under estimated:
MLE -0.002 (\(\pm\) 0.022); ELBn -0.039 (\(\pm\) 0.062); and NA -0.081
(\(\pm\) 0.061).

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods.}
\end{figure}

Interestingly, there was an interactive effect of the estimate accuracy
between sample size and the value of \(\lambda\). All methods were more
accurate with larger sample sizes, and smaller values of \(\lambda\).

\begin{figure}
\centering
\includegraphics{figures/n_vary_est_b.png}
\caption{Distribution of size spectra parameter estimates. Vertical line
is the known parameter wich describes the bounded power law distribution
from which the body size estimates were sampled. As n increases (top to
bottom) and \(\lambda\) increases (left to right), the accuracy of the
estimate improves across all methods.}
\end{figure}

@@

All of the methods performed similarly when there was no relationship
across the hypothetical gradient. The Type I error rate for MLE, ELBn
and NAS were 5\%, 5.7\% and 4.6\%, respectively. However, the confidence
intervals for the binning methods were \textasciitilde3 times as wide as
the MLE method (mean CI widths: MLE = 0.116, ELBn = 0.372; NAS = 0.332).

Of the relationships which were estimated to be significant, the MLE,
ELBn, and NAS method had a 58\%, 49.1\% and 54.3\% probability of
indicating a negative relationship, respectively.

\begin{figure}
\centering
\includegraphics{figures/static_b_lambda_spaghetti.png}
\caption{Individual regression estimates when no relationship exists
across the hypothetical environmental gradient. The methods are
separated by rows, and the left and right column show relationships
which were non-significant and significant, respectively. The individual
regressions coefficients are colored-coded to indicate positive and
negative relationship estimates.}
\end{figure}

@@

The performance of all methods declined when trying to detect variation
in the \(\lambda\) parameter between -1.9 to -2.1 across the
environmental gradient. The mean \(\beta_1\) coefficient estimates for
all methods were closely associated with the known relationship. Once
again the variation in the estimates using the MLE method were much
smaller that the estimates using both of the binning methods (Figure 4).

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_small_relationship_density.png}
\caption{Mean difference between relationship estimate and known
relationship coefficient for each method. Here, \(\lambda\) varied from
-1.9 to -2.1, with a kown relationaship of -0.1 across the gradient.}
\end{figure}

Of the 1000 simulation replicates, the MLE only detected a significant
relationship 90\% of the time, whereas the ELBn and NAS methods detected
significant relationships 19 and 23\% of the time, respectively. Of the
replicates which were significant, the confidence interval contained the
true value of the relationship 94.6, 88.4, and 89.1\% of the time for
the MLE, ELBn, and NAS methods, respectively.

\begin{figure}
\centering
\includegraphics{figures/small_lambda_spaghetti.png}
\caption{Linear regressions across the hypothetical gradient when lambda
varies from -1.9 to -2.1. Each method is displayed in the rows, and the
left colum shows the non-significant estimates, while the right column
shows the relationships which were signficant (\(\beta_1\) coefficient
p-value \textless{} 0.05). The colors indicate when the beta coefficient
estimate was negative (pink) and when it was positive (purple).}
\end{figure}

Once again, the width of the confidence intervals for both of the
binning methods were \(\ge\) 2.8 times that of the MLE method (mean
\(\pm\) SD MLE CI width: 0.117 \(\pm\) 0.051; ELBn CI width: 0.378
\(\pm\) 0.169; NAS CI width: 0.332 \(\pm\) 0.150).

On average, the relationship estimates across the gradient were similar
to the know value (mean deviation: MLE = 0.00001, ELBn = -0.009, NAS =
-0.012). However, when only looking at the relationships which were
significant, this variation increased considerably(MLE = -0.002, ELBN =
-0.052, NAS = 0.034).

@@

For both empirical datasets, the direction and magnitude of chnage
(i.e.~\(\beta_1\) coefficients) are generally in agreement. Size spectra
parameters consistently increase (become flatter) in the AMD data (Fig.
XX A). Likewise, the size spectra parameters consistently increase
(become steeper) with increasing temperature across the NEON sites (Fig
XX D)

\textbf{NOTE} combine all the figures below into one with 4 panels (AMD
= A, B; NEON = C, D)

\begin{figure}
\centering
\includegraphics{figures/AMD_plot.png}
\caption{A}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/amd_relationship.png}
\caption{B.Relationship estimates across AMD gradient}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/NEON_plot.png}
\caption{C.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/neon_relationship.png}
\caption{D.}
\end{figure}

Because the \(\beta_1\) coefficient estimates are similar, and the range
of the gradient in the AMD data is relatively small (9.5), the absolute
change in the size spectra parameter across the AMD gradient are similar
regardless of method used (range: 0.59 to 0.72). Likewise, the absolute
change in size spectra parameters across the NEON temperature gradient
ranges from 0.06 to 0.165, depending on the method used.

\emph{alternative} absolute difference of \textasciitilde0.1 units in
\(\lambda\) across gradients is similar to differences reported in other
studies (seasonally: 0.16, Mcgarvey and Kirk; landuse: 0.15 Martinez;
O'Gorman 2017 \textasciitilde0.25, Yvon and Dossena papers
\textasciitilde0.1? {[}need to redownload some of these and
compare{]})(Dossena et al. 2012, O'Gorman et al. 2017, McGarvey and Kirk
2018)

@@

Estimates of the relationship between lambda and a hypothetical gradient
varied depending on the method used. However, interpretations of
empirical data were broadly consistent across methodologies. Upon closer
inspection, the values of size spectra parameters in the empirical data
were considerably shallower than -2. Given that we found the performance
of all methods increased with shallower \(\lambda\)'s we wanted to
investigate how the methods performed with simulated values more similar
to the empirical estimates of parameters describing size spectra
relationships. Therefore, we repeated the simulation process as in the
main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5.

@@

We found generally less variability in the relationship estimates across
a gradient of distributions described with shallower \(\lambda\)
parameters. All of the relationship coefficient estimates
(\(\beta_1\))'s across all replicates were significant regardless of
method used. The confidence interval for \(\beta_1\) estimate for the
MLE method contained the true value 94.7\% of the time. However, the
true value was in the confidence interval for the ELBN and NAS method
only 90.9\% and 76.9\% of the time, respectively. Once again, the mean
width of the CI for the binning methods were \textasciitilde2 times as
large as for the MLE method. Mean \(\pm\) SD CI width: MLE = 0.0442
\(\pm\) 0.0196; ELBn = 0.0920 \(\pm\) 0.0482, NAS = 0.0798 \(\pm\)
0.0324.

On average, the relationship across the gradient was under estimated,
but by nearly an order of magnitude less than when simulating across
steeper values of \(\lambda\). Differences between known and estimated
relationship coefficients: MLE -0.0007 (\(\pm\) 0.0079); ELBn -0.0058
(\(\pm\) 0.0177); NAS -0.0251 (\(\pm\) 0.0124).

\begin{figure}
\centering
\includegraphics{figures/shallow_lambda_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods.}
\end{figure}

@@

Measuring parameters describing the decline in abundance with increasing
body size in communities is being done with increasing frequency across
ecology. Previous work has investigated the accuracy and inherent biases
associated with different estimation methods. However, how these
innaccuracies and biases compound across environmental gradients remains
uncertain, making it difficult to detecting variation in size spectra
parameters across environmental gradients with confidence. Here, we
sampled body sizes from known distributions and estimated their
parameters across a hypothetical gradient in order to assess the
relationship coefficients obtained.

Generally, there was large agreement in the direction and magnitude of
average relationship coefficient estimates across methods. The estimate
from the MLE method was always closer to the known relationship, and
always had the smallest variation around the estimate.

All of the methods performed equally well when the \(\lambda\) value did
not vary across the hypothetical gradient, with a Type I error rate of
\textasciitilde5\%, exactly as would be expected under the assumptions
of frequentist statistics. This indicates that previous studies which
have detected no difference in size spectra parameters across sites can
be interpreted with confidence, regardless of the method used.

The performance of all methods improved as sample size increases, and as
the exponent of the size spectra relationship gets larger (more
shallow). As either or both of these variables change, the difference in
estimated relationship coefficients declines across the methods. This is
particularly interesting given the fact that both empirical data sets of
stream communities examined here are shallower than the expected
relationship of \(\lambda = -2\), and the conclusions of the change in
size spectra relationships across the gradients are not dependent on
method used. If communities under study are in fact described by
shallower exponents, the method used may not be critical in the
conclusions reached. However, the MLE method performs as well or better
than the two binning methods examined here under all contexts and should
be the preferred method moving forward. At a minimum, future studies
should report MLE estimates to ensure that the results are not dependent
on the methodology used.

Alarmingly, the performance of all methods declined dramatically when
trying to detect smaller variation in \(\lambda\) values. We simulated
observations from distributions varying by an absolute value of 0.2
units. This range of variation is similar to that which has been
reported in many seminal works on variation in size spectra parameters.
When using the MLE method, significant relationships were only detected
\textasciitilde88\% of the time, whereas both of the binning methods
only detected significant trends \textasciitilde20\% of the time. Of the
simulation replicates which were significant, nearly all of them were
negative across all methods except for one single replicate using the
NAS method which was positive (Figure. 5).

Despite the drop in performance with reduced variation in \(\lambda\)
values, when a significant relationship was observed, it was generally
in the correct direction and of a similar magnitude. This suggests that
previously reported significant changes in size spectra parameters
across environmental gradients and in experimental manipulations are
plausible, and the magnitude of the relative change is a reasonable
estimate. Given that all of the data within a study is treated
identically, the the over all chnage in size spectra parameters is
likely reasonable. However, the biases and inconsistencies in
relationship estimates presented here suggest that it would be difficult
if not impossible to compare the relative changes across different
published studies which use different methods. Publication of individual
body size data with future studies of size spectra relationships would
greatly aid in our ability to generalize changes to this fundamental
aspect of community organization across spatiotemporal scales and in
response to environmental stressors and perturbations.

@@

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-dossena2012}{}}%
Dossena, M., G. Yvon-Durocher, J. Grey, J. M. Montoya, D. M. Perkins, M.
Trimmer, and G. Woodward. 2012.
\href{https://doi.org/10.1098/rspb.2012.0394}{Warming alters community
size structure and ecosystem functioning}. Proceedings of the Royal
Society B: Biological Sciences 279:3011--3019.

\leavevmode\vadjust pre{\hypertarget{ref-edwards2020}{}}%
Edwards, A. M., J. P. W. Robinson, J. L. Blanchard, J. K. Baum, and M.
J. Plank. 2020. \href{https://doi.org/10.3354/meps13230}{Accounting for
the bin structure of data removes bias when fitting size spectra}.
Marine Ecology Progress Series 636:19--33.

@@

\leavevmode\vadjust pre{\hypertarget{ref-mcgarvey2018}{}}%
McGarvey, D. J., and A. J. Kirk. 2018.
\href{https://doi.org/10.1007/s10750-017-3448-0}{Seasonal comparison of
community-level size-spectra in southern coalfield streams of {West
Virginia} ({USA})}. Hydrobiologia 809:65--77.

@@

>>>>

\title{Manuscript}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Jeff S. Wesner\textsuperscript{4}}
\date{04 September, 2022}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\setstretch{1}
\textsuperscript{1} Colorado Mesa University, Grand Junction, CO, USA\\
\textsuperscript{2} Great Lakes Research Center, Michigan Technological
University, Houghton, MI USA\\
\textsuperscript{3} Louisiana Universities Marine Consortium, Chauvin,
LA USA\\
\textsuperscript{4} Dept. of Biology, University of South Dakota,
Vermillion, SD, USA

@@

Body size distributions are a fundamental characteristic of communities.
The remarkable consistency of these relationships across spatiotemporal
scales and habitat types has led them to be recommended as a
``universal'' indicator of ecological status (Petchey and Belgrano
2010)(Petchey and Belgrano 2010). Size metrics have commonly been used
in marine systems, and are increasingly being applied to assess the
condition of freshwater ecosystems (Martinez et al.~2016, Pomeranz
2018)(Pomeranz et al. 2019, \textbf{martinez2016?}).

@@

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
on empirical body sizes of stream benthic communities reported in
Pomeranz et al.~(2020)(Pomeranz et al. 2022). Our results are not
dependent on the range of body sizes, and we show the results of other
body size ranges in the supplemental information.

For the main analysis, presented here, we sampled \(n = 1000\) body
sizes from distributions described by five different \(\lambda\)'s:
(-1.5, -1.75, -2.00, -2.25, -2.5), such that our model for \(\lambda\)
is:

\[\lambda = Norm(\alpha -0.5X, \sigma)\]

Each value of \(\lambda\) was assumed to come from a community across a
hypothetical environmental gradient \(X\). Values of \(X\) for each
community are uniformly distributed from -1 to 1. We repeated the data
simulation 1000 times (reps). The main results presented here were not
dependent on the range of \emph{x}-values or the number of sites
(supplemental information).

\hypertarget{sample-size-n}{%
\subsubsection{\texorpdfstring{Sample size,
\(n\)}{Sample size, n}}\label{sample-size-n}}

The number of observations in our simulations may bias the results.
Therefore, we repeated the simulations described above, but varied the
sample size \(n\). We tested values of
\(n = 200, 500, 1000, 5000, 10 000\).

@@

It is possible that our sampling or simulation framework could bias the
results obtained. In order to assess whether or not our simulations are
robust to this, we performed independent random samples from a bounded
power law for five communities across a hypothetical gradient as
described above, but we set \(\lambda = -2\) for all five communities.
Hence, we are simulating a scenario where the size spectra relationship
is invariant to our hypothetical environmental gradient, and we should
expect the estimated relationship coefficient \(\beta_1 = 0\).

@@

For the main analysis, we were interested in determining whether or not
the methods were able to detect relatively coarse changes to size
spectra parameters, specifically, a change of 0.5 units across each
value of \(X\), or an absolute change in \(\lambda\) values of 1.0
units. However, many seminal works on variation in size spectra
relationships in freshwater are on a much finer scale (on the order 0f
0.1 to 0.25 units). Therefore, we performed simulations as described
above, but varied \(\lambda\) from -1.9 to -2.1 across the hypothetical
environmental gradient. This equates to a know relationship coefficient
of \(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
change of 0.2 units.

@@

For each sample of \(M\) and each of the 1000 replicates, we estimated
the exponent \(\lambda\) using MLE methods modified from the the
\texttt{sizeSpectra} package (Edwards et al.~2020). In addition, we use
two common binning methods to estimate the OLS slope parameter in
log-log space. For the first binning method, we created 6 equal
logarithmic bins covering the range of body sizes. The count in each bin
was normalized by dividing by the bin width. This method has been used
by (refs here, I think a lot of Woodward's group uses this\ldots)
Throughout the manuscript, the normalized equal logarithmic binning
method will be referred to as ELBn. The second method was similar to
ELBn, but bins of Log\textsubscript{2} widths are used. The count in
each bin is normalized in the same way. This method has been used by
(Pomeranz et al. 2019, McGarvey et al. 2019) and is referred to as the
Normalized Abundance Spectrum (NAS). Previous work has investigated the
biases in \(\lambda\) estimates when using different methods (Edwards et
al.~(2017)(2017), Edwards et al.~2020(2020), White et al.~2008(2007)).
The focus of the present work is to investigate biases when estimating
the relationship or relative change in parameters across an
environmental gradient. Hence, we do not discuss the \(\lambda\)
estimates in the main text, but the results are presented in the
Supplmental Information.

@@

For each simulation replicate, we estimated how the size spectra
parameters varied across the hypothetical gradient. Simple OLS
regression were conducted in the form
\(\lambda_{estimate} ~ \beta_0 + \beta_1 * X\). The distribution of the
relationship coefficient, \(\beta_1\), were plotted compared to the
known relationship.

@@

We re-analyze two data sets of benthic macroinvertebrate communities
from stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage (AMD) stress gradient. Details of the
sample collection and processing can be found in (Pomeranz et al. 2019).
Briefly, all individuals from each sample were identified to the lowest
practical taxonomic unit and body lengths were measured using image
processing software from photos taken with a camera mounted to a
dissecting microscope. Body mass was estimated using taxon-specific
published length weight regressions.

The second data set was from the wadeable stream sites of the National
Ecological Observatory Network (National Ecological Observatory Network
(NEON) 2022). NEON stream sites are located across a wide temperature
gradient in the United States, from Puerto Rico to Alaska. Quantitative
macroinvertebrate samples were collected using the most appropriate
method based on the local habitat. All individuals were identified and
had their body lengths measured, and body mass was estimated using
published length weight regressions. This data has been analyzed
previously using size spectra methods as described in (Pomeranz et al.
2022). Detailed methods of the sampling collection and data processing
methods can be found on the NEON website
({[}\url{https://www.neonscience.org/data-collection/macroinvertebrates}{]},
or macroinvertebrate DPI pubs). Coefficient estimates (\(\pm\) 1
standard deviation) are compared across methods. This allows us to
determine whether or not the main results would differ depending on the
method used.

@@

Alternate plot version

\begin{figure}
\centering
\includegraphics{figures/PLB_main_v2.png}
\caption{alternate plot version, thoughts on which you prefer??}
\end{figure}

All methods performed reasonably well in detecting the known
relationship across the hypothetical environmental gradient (mean
\(\beta_1\) estimates ranged from -0.419 to -0.498). However, the SD for
the NAS and ELBn methods were 2-3x larger that that of the MLE method
(Figure \textsubscript{XX}). This result was consistent across
simulation parameters, including sample size, number of sites, and
\(\lambda\) values.

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_relationship_density.png}
\caption{Halfeye plots for the estimated relationship coefficient
estimates across 1000 replicates and three different methods. The point
below each distribution plot is the median estimate, the thick and thin
bars represent the 66th and 95th percentile, respectively. Vertical line
is the known relationship. The median estimate for the MLE method was
closest to the known relationship, and had the narrowest distribution
across replicates.}
\end{figure}

\begin{verbatim}
##   Model   p50   p25  p975
## 1   MLE -0.49 -0.50 -0.48
## 2  ELBn -0.49 -0.52 -0.29
## 3   NAS -0.44 -0.46 -0.30
\end{verbatim}

@@

All of the methods performed similarly when there was no relationship
across the hypothetical gradient. The mean \(\beta_1\) estimates for all
three methods ranged from -0.0005 to 0.0002. However, the distribution
of coefficient estimates was much wider for the two binning methods; the
95th percentile of the MLE distributions is narrower thatn the 66th
percentiles for the NAS and ELBn methods.

\begin{figure}
\centering
\includegraphics{figures/PLB_static_b_relationship_density.png}
\caption{Individual regression estimates when no relationship exists
across the hypothetical environmental gradient. The methods are
separated by rows. The median for all methocs is approximately zero, but
the distribution for the MLE method is narrower than the two binning
methods.}
\end{figure}

\begin{verbatim}
##   Model p50   p25 p975
## 1   MLE   0 -0.01 0.04
## 2   NAS   0 -0.04 0.12
## 3  ELBn   0 -0.05 0.13
\end{verbatim}

@@

The performance of all methods to detect small variation in the
\(\lambda\) parameter (between -1.9 to -2.1) across the environmental
gradient was also investigated. The median \(\beta_1\) coefficient
estimates for the MLE method was -0.1, which exactly matched the value
for the hypothetical relationship, while the other two methods both had
median estimates of 0.9. The upper 97.5 percentile of the distribution
was below 0 for the MLE method, meaning that most of the estimates had
the correct sign. In contrast, the upper 97.5 percentile for both the
NAS and ELBn methods were greater than 0 at 0.027 and 0.052,
respectively. (Figure 4).

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_small_relationship_density.png}
\caption{Distribution of the estimated relationship for each method.
Here, \(\lambda\) varied from -1.9 to -2.1, with a known relationaship
of -0.1 across the gradient indicated by the dashed vertical line. The
distribution of relationship estimates was approximately centered at the
known relationship for all methods, with a narrower distribution
observed for the MLE estimates.}
\end{figure}

\begin{verbatim}
##   Model   p50   p25  p975
## 1   MLE -0.10 -0.11 -0.06
## 2   NAS -0.09 -0.13  0.03
## 3  ELBn -0.09 -0.13  0.05
\end{verbatim}

@@

For both empirical data sets, the direction and magnitude of change
(i.e.~\(\beta_1\) coefficients) are generally in agreement. Size spectra
parameters consistently increase (become flatter) in the AMD data (Fig.
XX A). Likewise, the size spectra parameters consistently increase
(become steeper) with increasing temperature across the NEON sites (Fig
XX D)

\textbf{NOTE} I need to re-plot these in ggplot and and stitich
together. Currently stitching pre-made png files, I think that's why it
looks so funky.

\begin{figure}
\centering
\includegraphics{figures/empirical_combined.png}
\caption{Figure of empirical data estimates. All of the methods estimate
the same sign of the relationship, but the estimates from the binning
methods are generally greater than the MLE estimates.}
\end{figure}

The \(\beta_1\) coefficient estimates for the AMD data are all positive
and of relatively similar magnitudes across methods. The MLE estimate
has the smallest magnitude and SD (0.063 \(\pm\) 0.011 SD), The ELBn
method is slightly larger and has a wider SD : (0.072 \(\pm\) 0.018).
Finally, the NAS method has the largest estimate and an intermediate
standard deviation (0.076 \(\pm\) 0.013). The range of the gradient in
the AMD data is relatively small (9.5), meaning that the absolute change
in the size spectra parameter (\(\lambda\)) across the AMD gradient are
similar regardless of method used (range of absolute: 0.59 to 0.72).

Similar to the AMD results, the \(\beta_1\) coefficient estimates for
the NEON data are all negative and have similar magnitudes across
methods. The MLE estimate is once again the smallest (-0.006 \(\pm\)
0.001 SD). The NAS and ELBn methods are larger at -0.004 \(\pm\) 0.002,
and -0.002 \(\pm\) 0.002, respectively. Although the estimates are
similar and have a small magnitude, the range of the gradient in this
data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
change in \(\lambda\) estimates ranges from 0.06 to 0.17, depending on
the method used.

@@

Estimates of the relationship between lambda and a hypothetical gradient
varied depending on the method used. However, interpretations of
empirical data were broadly consistent across methodologies. Upon closer
inspection, the values of size spectra parameters in the empirical data
were considerably shallower than -2. Given that we found the performance
of all methods increased with shallower \(\lambda\)'s we wanted to
investigate how the methods performed with simulated values closer to
the empirical estimates of parameters describing size spectra
relationships. Therefore, we repeated the simulation process as in the
main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5,
with a known relationship of \(\beta_1 = -0.2\)

@@

We found generally less variability in the relationship estimates
(\(\beta_1\)) across a gradient of distributions described with
shallower \(\lambda\) parameters. However, the median MLE estimate was
once again closest to the hypothetical relationship and had the
narrowest distribution. The median ELBn estimate was close to the
hypothetical relationship but had a slightly wider distribution.
Finally, 95\% of the NAS estimates were greater than the hypothetical
relationship. In general, the distributions from the different
methodologies followed the same overall pattern, with the MLE estimates
being the smallest, and the binning methods being larger (i.e., more
positive).

\begin{figure}
\centering
\includegraphics{figures/PLB_shallow_lambda_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods.}
\end{figure}

\begin{verbatim}
##   Model   p50   p25  p975
## 1   MLE -0.20 -0.21 -0.18
## 2  ELBn -0.19 -0.21 -0.16
## 3   NAS -0.17 -0.18 -0.15
\end{verbatim}

@@

Measuring parameters describing the decline in abundance with increasing
body size in communities is being done with increasing frequency across
ecology. Previous work has investigated the accuracy and inherent biases
associated with different estimation methods. However, how these
inaccuracies and biases compound across environmental gradients remains
uncertain, making it difficult to detect variation in size spectra
parameters across environmental gradients with confidence. Here, we
sampled body sizes from known distributions with varying parameters
(\(\lambda\)) and estimated the coefficient of the relationship
(\(\beta_1\)) across a hypothetical gradient in order to assess how the
results would vary depnding on the method used. Likewise, we compared
how the interpretation of previously published results could change
depending on the methodology used to estimate the relationship.

Generally, the estimate from the MLE method was always closer to the
known relationship in simulated data, and always had the smallest
variation around the estimate.

Binning methods are easy to use and interpret, which most likely
accounts for their wide use in ecological studies. However, aggregating
individuals into logarithmic bins removes a large source of the
variation within the data by collapsing body size variation into a
single values within bins. For example, all individuals placed into a
bin that ranges from 2-4 grams of mass are all treated as having a mass
of 3 grams, the midpoint of that bin. Likewise, a single abundance value
is taken for this bin, despite that fact that there is almost certainly
variation in the abundance of individuals that weigh \textasciitilde2
grams or \textasciitilde4 grams. By homogenizing the data in this way,
it seems that binning methods will always produce noisier results than
MLE from the binning process alone, since binning is akin to deleting
information that the model could otherwise use.

One of the benefits of using MLE is that all data points are retained
for use within the model.

The performance of all methods improved as sample size increases, and as
the exponent of the size spectra relationship gets larger (more
shallow). As either or both of these variables change, the difference in
estimated relationship coefficients declines across the methods. This is
particularly interesting given the fact that both empirical data sets of
stream communities examined here are shallower than the expected
relationship of \(\lambda = -2\), and the conclusions of the change in
size spectra relationships across the gradients are not dependent on
method used. If communities under study are in fact described by
shallower exponents, the method used may not be critical in the
conclusions reached. However, the MLE method performs as well or better
than the two binning methods examined here under all contexts and should
be the preferred method in future studies. At a minimum, future studies
should report MLE estimates to ensure that the results are not dependent
on the methodology used.

Despite the drop in performance with reduced variation in \(\lambda\)
values, when a significant relationship was observed, it was generally
in the correct direction and of a similar magnitude. This suggests that
previously reported significant changes in size spectra parameters
across environmental gradients and in experimental manipulations are
plausible, and the magnitude of the relative change is a reasonable
estimate. Given that all of the data within a study is treated
identically, the the over all change in size spectra parameters is
likely reasonable. However, the biases and inconsistencies in
relationship estimates presented here suggest that it would be difficult
if not impossible to compare the relative changes across different
published studies which use different methods. Publication of individual
body size data with future studies of size spectra relationships would
greatly aid in our ability to generalize changes to this fundamental
aspect of community organization across spatiotemporal scales and in
response to environmental stressors and perturbations.

@@

Both of the binning methods can perform as well as MLE under certain
circumstances (large sample size, shallow \(\lambda\)). However, it
would be extremely difficult to know if this was the case beforehand.
Furthermore, in order to confirm this result, MLE methods would likely
need to be employed. With the publication of the sizespectra package
(2017), producing MLE estimates of size spectra parameters is a
relatively easy task. Therefore, we recommend using this in all future
studies of size spectra relationships.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-edwards2020}{}}%
Edwards, A. M., J. P. W. Robinson, J. L. Blanchard, J. K. Baum, and M.
J. Plank. 2020. \href{https://doi.org/10.3354/meps13230}{Accounting for
the bin structure of data removes bias when fitting size spectra}.
Marine Ecology Progress Series 636:19--33.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-pomeranz2019}{}}%
Pomeranz, J. P. F., H. J. Warburton, and J. S. Harding. 2019.
\href{https://doi.org/10.1111/fwb.13196}{Anthropogenic mining alters
macroinvertebrate size spectra in streams}. Freshwater Biology
64:81--92.
